Notes for Classification of topological quantum matter with symmetries

### 2016-11-18 Latex packages/latex examples.tex
Read a bit about characters in S. Lang's language. I should learn more about modules to understand the idea of simple module and simple ring.
Fix some issue about latex package "physicist-taper". Add some crude simple notes about the classification paper.

\maketitle
\abstract{
Nothing here.
}
\tableofcontents
\begin{center}
    Bad diagrams
\end{center}
$$
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"]\ar[drr, phantom, "\#", very near
        start] & & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"]\ar[drr, phantom, "\#"] & & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
$$
\begin{center}
    A good diagram
\end{center}
$$
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"] &
        \ar[d,phantom, "\#", near start] & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
$$

\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    %\bibitem{book} 
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

\maketitle
\abstract{
Nothing here.
}
\tableofcontents
\section{Nomenclature}

\nomenclature{aa}{aaa}

\section{Diagram}

\begin{center}
    Bad diagrams
\end{center}
$$
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"]\ar[drr, phantom, "\#", very near
        start] & & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"]\ar[drr, phantom, "\#"] & & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
$$
\begin{center}
    Good diagrams
\end{center}
$$
    \begin{tikzcd}[column sep=small]
        E \ar[dr,"f"]\ar[rr,"\rho"] &
        \ar[d,phantom, "\#", near start] & F\ar[dl,"\rho'"] \\
        & G &  \,
    \end{tikzcd}
$$
$$ \begin{tikzcd}[column sep=small]
& G &  \, \\
E \ar[ur,"f"]\ar[rr,"\rho" below=3] &
\ar[u,phantom, "\#"] 
& F\ar[ul,"\rho'" right=3]
\end{tikzcd} $$

\section{Table}

\begin{table}[ht]
	\centering
	\caption{Classification}
	\label{tab:classification}
	\vspace{+0.5pt}
	\rowcolors{1}{white}{lightgrey}
	\tabulinesep=1.2mm
	\begin{tabu}{  X[l] X[l] X[l] X[l] }
		\textit{Symmetry} & \textit{Spatial Dimension} &
		\textit{Result} & \textit{Other Keywords} \\
		\hline
		T & 0 & An intger: the number of particle-occupied Kramers
		doublet states & \\
		T & 1 & None & \\
		T & 2 & $\mathbb{Z}_2$ & \\
		T & 3 & $\mathbb{Z}_2$ & $3D$ crystals have additional
		$3\mathbb{Z}_2$ invariant$ \Rightarrow$ "weak topological
		insulators \\
		Q(?) & 2 & Characterized by $\mu$ in units of $e^2/h$ & TKNN \\
		Q(?) & even  d & Topological invariant ($k$-th Chern number) \\
		Q(?) & 0 &  number of single-particle states with negative
		energy ($E< E_F = 0$), which are filled with electrons. \\
		T\& Q \\
		No T \& No Q & 0 & $\mathbb{Z}_2$ \\
		No T \& No Q & 1 &$\mathbb{Z}_2$ &"majorana chain"  \\
		No T \& No Q & 2 & Topological number is integer. & Even-odd
		effects. \\
		\bottomrule
	\end{tabu}
\end{table}
\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{book}  s
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

### 2016-11-18 Notes of various papers/Classification of topological quantum matter with symmetries. Rev.M.P/notes.tex
Fix some issue about latex package "physicist-taper". Add some crude simple notes about the classification paper.

\maketitle
\abstract{
As title suggests.
}
\tableofcontents
\paragraph{Asks}
\begin{enumerate}
    \item Page 6. Why the scalar in Schur's lemma becomes of unit lenght.
    \item page 9. What does it mean by:
        \begin{quote}
            Note that unitary symmetries, which commute with the
            Hamiltonian, allow us to bring the Hamiltonian into a
            block diagonal form.
        \end{quote}
    \item page 9, table I. When he talks about \textit{codimension},
        what is the dimension of the whole space? ($3$? $3+1$?).
        Similar problem also exists in page 10.
        Note, he mentions codimension of gaples modes on page 11. He
        asserts that codimension $1$ means $1$ dimension less than the
        bulk. But it should be strange to compare the dimension of
        defects with the dimension of the bulk.

@@

\paragraph{Revisit}
\begin{enumerate}
    \item page 12, bottom. How this procedure of relating real and
        complex classification is done?
\end{enumerate}
\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    %\bibitem{book} 
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

\maketitle
\abstract{
As title suggests.
}
\tableofcontents
\section{Doubts}
\label{sec:Doubts}

\paragraph{Asks}
\begin{enumerate}
    \item Page 6. Why the scalar in Schur's lemma becomes of unit lenght.
    \item page 9. What does it mean by:
        \begin{quote}
            Note that unitary symmetries, which commute with the
            Hamiltonian, allow us to bring the Hamiltonian into a
            block diagonal form.
        \end{quote}
    \item page 9, table I. When he talks about \textit{codimension},
        what is the dimension of the whole space? ($3$? $3+1$?).
        Similar problem also exists in page 10.
        Note, he mentions codimension of gaples modes on page 11. He
        asserts that codimension $1$ means $1$ dimension less than the
        bulk. But it should be strange to compare the dimension of
        defects with the dimension of the bulk.

@@

\paragraph{Revisit}
\begin{enumerate}
    \item page 12, bottom. How this procedure of relating real and
        complex classification is done?
\end{enumerate}
\section{Summary}
\label{sec:Summary}

\subsection{About topological matter}
\label{sec:About-topological-matter}
Topological states:
\begin{itemize}
    \item Topological insulators cannot be distinguished from
        ordinary, topologically trivial insulators in terms of their
        symmetries.
    \item topological insulators' topological nontriviality cannot be
        detected by a local order parameter.
    \item insulating electronic band structures can be categorized in
        terms of topology.
    \begin{itemize}
        \item time-reversal are crucial in makeing a spin-orbit induced
            topological insulators. $ \Rightarrow$ symmetry-protected
            topological phases (\nomen{SPT}).
        \item symmetry-protected topological phase can also arise from
            spatial symmetry.
    \end{itemize}
    \item There is a direct analogy between \nomen{TI}s and
        \nomen{TSC}s (topological insulators and topological
        superconductors).
    \item \nomen{bulk-boundary correspondence} $ \Leftarrow$
        topological phase transition. (pp. 11).
    \item \nomen{nodal systems} can exhibit band topology even though
        their bulk gap closes at certain points in BZ.
    \item All of the above are understood in terms of non-interacting
        or mean-field Hamiltonians. Less is known about strongly
        correlated systems.
\end{itemize}

\subsection{Features of Tenfold for Gapped System}
\label{sec:Tenfold-for-Gapped-System}
\nomen{Topological equivalence}. $l:$ adiabatic path or continuous path in
phase diagram. Then topological class are characterized by invariance
of gap under $l$.

\nomen{Trivial class}: The topological classes of state
characterized by atomic insulators.

\nomen{Weak topological phase}: nontrivial topological phase relies
on lattice-translational symmetry.
\nomen{Strong topological phase}: ... that does not rely on
translational symmetry.

\paragraph{Classfication of Gapless}
\begin{itemize}
    \item $2$ complex symmetry classes, $8$ real symmetry classes.
    \item The presence or absence of weak TIs or TSCs in a given
        symmetry classes can be deduced from that of strong TIs or
        TSCs in lower dimension in the same symmetry class. So the
        classification aims for the strong ones.
        % TODO expands following in the future.
    \item \nomen{dimensional shift} (pp. 11).
    \item \nomen{primary series}, \nomen{first/second decendants},
        \nomen{even series}:
        class of items in the periodic table. (pp. 12).
\end{itemize}

\paragraph{Classfication with defects}
\begin{itemize}
    \item \nomen{defect Hamiltonian} $H(k,r)$,$r=(x,t)$. Key: $r$
        varies slowly ($\xi|\grad_r H(r,k)|\ll \varepsilon_g$). page
        13.
    \item Characterization (wrapped by $S^D$)
        \begin{enumerate}
            \item $s$: AZ symmetry class
            \item $d$: bulk dimension
            \item $d_c$: defect codimension
            \begin{itemize}
                \item real AZ class: $s-\delta$ (mod $8$)
                    ($\delta=d-D$, topological dimension)
                \item complex AZ classes, similarly
            \end{itemize}
        \end{enumerate}
    \item Relate real and complex: forgetful functor.
    \item \nomen{bulk-defect correspondence}: guarantees gapless
        defect excitation.
\end{itemize}
\paragraph{Topological Invariants} $ $

\nomen{Base space}:
\begin{equation}
\begin{tikzcd}[]
    (\vb{k},\vb{r})\in \text{BZ}^d\times \mathcal{M}^D
    \ar[rr, "\text{compactify}"] &\, & S^{d+D}
\end{tikzcd}
\end{equation}
Table 3 in page 13, is a summary of invariants:
\nomen{s}:= AZ symmetry class. \nomen{Ch}: Chern number.
\nomen{CS}:Chern-Simons invariant. \nomen{$\nu$}: Winding number.
\nomen{FK}: Fu-Kane.
\begin{table}[H]
    \centering
    \caption{caption}
    \label{tab:label}
    \begin{tabu}{c c c}
        $\,$ & $s$ even & $s$ odd \\
        \hline
        $\mathbb{Z}$ & Ch & $\nu$ \\
        $\mathbb{Z}_2^{(1)}$ & CS & FK \\
        $\mathbb{Z}_2^{(2)}$ & FK & CS ($\widetilde{\text{CS}}$) \\
        \bottomrule
    \end{tabu}
\end{table}

\subsection{Chern Number}
\label{sec:Chern-Number}
It detects an obstruction in defining a set of bloch wave functions
smooothly over the base space.

\nomen{Fibres} The fibres of the fibre-bundle on the base space are
wave functions under adiabatic changes. They may be twisted.

\nomen{Twist}: Think of the M\"obius strip as a twisted fibre bundle.
A good resource:
\href{http://math.stackexchange.com/questions/1377855/cylinder-and-m%C3%B6bius-strip-as-fiber-bundles-trivializations-and-cocycles}{Cylinder
and Möbius strip as fiber bundles}. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.2\linewidth]{pics/mobius-fibre.png}
    \caption{M\"obius Strip, 
        Credit:
        \href{https://www.quora.com/What-is-a-Fiber-Bundle-in-laymans-terms-What-is-it-used-for}{Quora} }
\end{figure}

\begin{thm}
    The set of Bloch functions (i.e. projectors) defines a map from the
    base space to $U(N_+ +N_-)/U(N_+)\times U(N_-)$. Topological
    distinct maps of this type can be classified by their homotopy
    group
    \begin{equation}
        \pi_{d+D}[U(N_+ +N_-)/U(N_+)\times U(N_-)]
    \end{equation}
\end{thm}

When $N_\pm$ is large enough, and when $d+D$ is even, this group is
$\mathbb{Z}$, the integer topological invariant is the Chern number.

\subsection{Winding number}
\label{sec:Winding-number}
\begin{itemize}
    \item Can only be defined in the presence of chiral symmetry.
    \item Relavant space: $U(N), \Rightarrow$
    \item $Q$ matrix (projector) Can be classified by
        $\pi_{d+D}[U(N)]$, which is nontrivial when $d+D$ is odd. When
        $d+D$ is odd, $\pi_{d+D}[U(N)]=\mathbb{Z}$, with maps
        characterized by the winding number.
\end{itemize}

\subsection{Chern-Simons invariant}
\label{sec:Chern-Simons-invariant}
\begin{itemize}
    \item Can be defined when $d+D$ is odd.
    \item Symmetry $\rightarrow$ Quantized/Discrete CS invariant. e.g.
        \begin{itemize}
            \item chiral symmetry quantizes CS
        \end{itemize}
    \item Not gauge-invariant.
        \begin{itemize}
            \item But $W:= \exp{2\pi i \,\text{CS}}$ is gauge-invariant.
        \end{itemize}
\end{itemize}

\subsection{Fu-Kane Invariant}
\label{sec:Fu-Kane-Invariant}
\begin{itemize}
    \item Construction requires special attention in the choice of basis
\end{itemize}

\subsection{K-theory Approach}
\label{sec:K-theory-Approach}
\begin{itemize}
    \item Topological nontrivial phases $\Rightarrow$ have massive
        Dirac Hamiltonian representatives
\end{itemize}
\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{review}
    \url{http://journals.aps.org/rmp/abstract/10.1103/RevModPhys.88.035005}
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}
